Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-29 introduction.tex
related works fixes

\section{Related work}
\secL{Related-work}
Many different design space exploration algorithms have been proposed
in literature.  What makes the exploration of embedded systems so
complex is not only the huge size of the parameter spaces involved,
but also the tight interaction between parameters in relation to their
effect on the objective to be optimised, which take the form of
\emph{parameter dependencies} (see~\defR{dependency}). The 
motivations of an exploration algorithm and how it deals with such
complexity are often rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

@@

Authors of \cite{dellnitz2005covering,dellnitz2009computation} propose
three different methods for the numerical approximation of the Pareto
set. The first two require the knowledge of the gradient of the
objective functions, which is usually not available in embedded system
design, since the closed form expression of the objective functions is
not known a priori. The third one, called ``Sampling algorithm'', does
not need this requirement and is the closest to ours, although the
mechanism is different. At every iteration a set of regions of
configuration space is constructed with the aim to cover the Pareto
set. The size of these regions becomes smaller at each iteration so
that the covering becomes tighter. A random set of test points are
evaluated in each new region; only the regions containing
non-dominated points will be considered in the next iteration. Our
approach is different. First, in the Sampling algorithm only the
selected regions are examined, even though there is no guarantee that
all the Pareto points are included in them. As a consequence, the
Sampling algorithm may erroneously and prematurely neglect some
regions that may possibly contain a non-negligible number of Pareto
points. Our algorithm avoids this, since even though a region has not
provided non-dominated points, it is not neglected, rather it is
simply populated with fewer test points, thus giving the chance to
find non-dominated points in them. From this behavior also comes
another important difference. The Sampling algorithm of
\cite{dellnitz2005covering} keeps reinforcing the analysis of the
regions selected in the previous iterations. In other words, it keeps
zooming into the regions that have been considered ``interesting'',
i.e. rich of non-dominated points, in the previous iterations. The
philosophy behind our algorithm is different. A region, which has been
considered interesting at a certain iterations, may be not in the
future, since it has been thoroughly analyzed. On the other hand, a
region, initially uninteresting, may turn out to be interesting in the
future, once all the other regions have been already exploited.
Therefore, from an iteration to the next, our algorithm zoom in on
regions that are currently interesting and zoom out from the
uninteresting ones,  spanning in a broader and more dynamical way the
configuration space. In addition, in the Sampling region only the
number of non-dominated points is accounted to see how interesting a
region is. On the contrary, we also consider the \emph{innovation}, as
already anticipated in the introduction.

Many studies, as \cite{coello_easmop} and others, focus on genetic
approaches to solve multiobjective optimization problems. Genetic
approaches have many advantages: they provide good accuracy, they are
customizable and very general and require no a-priori knowledge to the
designer. The strong point of genetic approaches can be summarized as
follows: they consist in an exploration that is sufficiently broad
(the mutation avoids being rigidly restricted to limited parts of
parameter space), and, at the same time, not too scattered as it is
guided by the performances of the already evaluated configurations.
Different variants of these algorithms have been proposed in
literature, including mixed approaches as in~\cite{Ascia2011382},
where authors specifically focuse on increasing the performance of a
evolutionaly exploration strategy, proposing a fuzzy based technique
to avoid the actual simulation of a certain number of configurations.
In particular, in this work we compare the proposed approach against a
well known Multi-Objective Genetic Algorithm~\cite{knowles_techrep06}
which is at best of our knowledge one of the most appropriate for the
particular scenario we adopted as testbed.  We claim that the approach
presented in this paper has most of the benefits of the genetic
approaches, although the rationale is completely different.

>>>>

\section{Related work and Our Contribution}
\secL{Related-work}
Many different design space exploration algorithms have been proposed
in literature.  What makes the exploration of embedded systems so
complex is not only the huge size of the parameter spaces involved,
but also the tight interaction between parameters in relation to their
effect on the objective to be optimised, which take the form of
\emph{parameter dependencies} (see~\defR{dependency}). The 
motivations of an exploration algorithm and how it deals with such
complexity are often rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

@@

Authors of \cite{dellnitz2005covering,dellnitz2009computation} propose
three different methods for the numerical approximation of the Pareto
set. The first two require the knowledge of the gradient of the
objective functions, which is usually not available in embedded system
design, since the closed form expression of the objective functions is
not known a priori. The third one, called ``Sampling algorithm'', does
not need this requirement and is the closest to the PS approach
presented in this work, although the
mechanism is different. At every iteration a set of regions of
configuration space is constructed with the aim to cover the Pareto
set. The size of these regions becomes smaller at each iteration so
that the covering becomes tighter. A random set of test points are
evaluated in each new region; only the regions containing
non-dominated points will be considered in the next iteration.
However, an important feature differentiates it from PS. The Sampling algorithm only the
selected regions are examined, even though there is no guarantee that
all the Pareto points are included in them. As a consequence, the
Sampling algorithm may erroneously and prematurely neglect some
regions that may possibly contain a non-negligible number of Pareto
points. The PS algorithm avoids this, since even though a region has not
provided non-dominated points, it is not neglected, rather it is
simply populated with fewer test points, thus giving the chance to
find non-dominated points in them. From this behavior also comes
another important difference. The Sampling algorithm of
\cite{dellnitz2005covering} keeps reinforcing the analysis of the
regions selected in the previous iterations. In other words, it keeps
zooming into the regions that have been considered ``interesting'',
i.e. rich of non-dominated points, in the previous iterations. The
philosophy behind our algorithm is different. A region, which has been
considered interesting at a certain iterations, may be not in the
future, since it has been thoroughly analyzed. On the other hand, a
region, initially uninteresting, may turn out to be interesting in the
future, once all the other regions have been already exploited.
Therefore, from an iteration to the next, our algorithm zoom in on
regions that are currently interesting and zoom out from the
uninteresting ones,  spanning in a broader and more dynamical way the
configuration space. In addition, in the Sampling region only the
number of non-dominated points is accounted to see how interesting a
region is. On the contrary, we also consider the \emph{innovation}, as
already anticipated in the introduction.

The most well documented and widespread class of algorithms adopted in the
design space exploration of embedded systems is probably the one of Evolutionary
Multiobjective Optimization Algorithms~\cite{coello_easmop}. For this
reason this is the class of algorithms that we will use for the
experiments discussed in~\secR{ee}.
Evolutionary approaches have many advantages: they provide good accuracy, they are
problem-independent and do not require any a-priori knowledge of the
system. The strong point of these approaches is that they consist in an exploration that is sufficiently broad
(the mutation avoids being rigidly restricted to limited parts of
parameter space), and, at the same time, not too scattered as it is
guided by the performances of the already evaluated configurations.
Different variants of these algorithms have been proposed in
literature, including mixed approaches as in~\cite{Ascia2011382},
where authors specifically focuse on increasing the performance of a
evolutionaly exploration strategy, proposing a fuzzy based technique
to avoid the actual simulation of a certain number of configurations.
In particular, in this work we will compare the proposed approach against a
well known SPEA2 multi-objective genetic algorithm~\cite{knowles_techrep06}
which is, at best of our knowledge, one of the most appropriate and
tested for the particular scenario we are assuming.  We
claim that the approach presented in this paper has most of the
benefits of the evolutionary approaches, although the rationale is
completely different.
